---
title: Off-Patent Drug Look-Up
subtitle: Ayush Noori
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
# knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```

# Dependencies

Load requisite packages and define directories.

```{r load-packages, message=FALSE, warning=FALSE}

# data manipulation
library(data.table)
library(purrr)
library(magrittr)

# excel manipulation
library(openxlsx)

# string manipulation
library(stringi)

# utilities
library(brainstorm)

```

Note that directories are relative to the R project path.

```{r define-directores}

# set directories
dir = c("C:/Users/ayush/OneDrive/Extracurricular/Entrepreneurship/gener8tor Tech Transfer/DTC Pharma/Off-Patent, Off-Exclusivity Drugs")
ddir = file.path(dir, "2021.07.20 - Drugs at FDA")

```

# Read Data

First, select dataset for analysis.

```{r select-dataset}

opt = c("Part I", "Part II", "Appendix") %>% purrr::set_names(.)
nm = opt[menu(opt, title = "Please select the desired dataset for analysis.")]

```

Read drugs which are off-patent and off-exclusivity.

```{r read-drugs}

# read data
drugs = as.data.table(read.xlsx(file.path(dir, "2021.06.14 - Off-Patent Drugs.xlsx"), sheet = nm, sep.names = " "))[, 1:3]

# make column names syntactically valid
colnames(drugs) = gsub(" ", "", colnames(drugs))

# create index column
drugs[, ApplNo := gsub("N0", "", ApprovedNDA, fixed = T)] %>%
  .[, ApplNo := gsub("N", "", ApplNo, fixed = T)] %>%
  .[, ApplNo := as.integer(ApplNo)]

```

Read data downloaded from [the Drugs\@FDA database](https://www.fda.gov/drugs/drug-approvals-and-databases/drugsfda-data-files). Note that while the online database is updated daily, the data files are updated once per week, on Tuesdays.

```{r read-fda}

# function to read data
rd = function(x) { fread(file.path(ddir, x)) }

# read database files
flist = list.files(ddir)
db = map(flist, rd)
names(db) = gsub("(.txt)|(.csv)", "", flist)

```

# Merge Data

Collapse products by pasting approved strengths.

```{r remove-duplicates}

# # function to paste duplicates
# paste_duplicates = function(x) {
#   rem_dup = map_dfc(x, ~paste(unique(.x), collapse = "; "))
#   return(rem_dup)
# }
# 
# # paste products
# products = db$Products[, !c("ProductNo", "ReferenceDrug")] %>%
#   .[, paste_duplicates(.SD), .SDcols = colnames(.), by = "ApplNo"] %>%
#   .[, -1]

```

Merge the drug list with the FDA datasets to create final table.

```{r merge-data}

# format marketing status
MarketingStatus = copy(db$MarketingStatus) %>%
  .[, ID := paste(ApplNo, ProductNo, sep = "_")] %>%
  .[, .(ID, MarketingStatusID)]

# format application docs
ApplicationDocs = copy(db$ApplicationDocs) %>%
  .[SubmissionType == "ORIG"] %>%
  .[, .SD[which.min(ApplicationDocsDate)], by = "ApplNo"] %>%
  .[, .(ApplNo, ApplicationDocsTypeID, ApplicationDocsDate, ApplicationDocsURL)]

# merge databases
mdrugs = merge(drugs, db$Products[, !c("ReferenceDrug")], by = "ApplNo", all.x = T, all.y = F) %>%
  .[, ID := paste(ApplNo, ProductNo, sep = "_")] %>%
  merge(MarketingStatus, by = "ID", all.x = T, all.y = F) %>%
  merge(db$Applications[, !c("ApplPublicNotes", "ApplType")], by = "ApplNo", all.x = T, all.y = F) %>%
  merge(ApplicationDocs, by = "ApplNo", all.x = T, all.y = F)

# rename columns
setnames(mdrugs, c("ApplNo", "ProductNo", "DrugName", "ApprovedNDA", "ActiveIngredient", "MarketingStatusID", "SponsorName", "ApplicationDocsTypeID", "ApplicationDocsDate", "ApplicationDocsURL"), c("Application", "Product", "Drug", "NDA", "ActiveIngredients", "Marketing", "Sponsor", "DocumentType", "DocumentDate", "DocumentURL"))

# reorder columns
mdrugs = mdrugs %>%
  .[, c("ID", "ActiveIngredients") := NULL] %>%
  .[order(Application, Product)] %>%
  setcolorder(c("Application", "Product", "Drug", "NDA", "Marketing", "Sponsor", "Form", "DosageForm", "Strength", "Ingredient"))

# map columns to factors
mdrugs %>%
  .[, Marketing := factor(Marketing, levels = db$MarketingStatus_Lookup$MarketingStatusID, labels = db$MarketingStatus_Lookup$MarketingStatusDescription)] %>%
  .[, DocumentType := factor(DocumentType, levels = db$ApplicationsDocsType_Lookup$ApplicationDocsType_Lookup_ID, labels = db$ApplicationsDocsType_Lookup$ApplicationDocsType_Lookup_Description)] 

# clean dosage form
mdrugs = mdrugs %>%
  .[, DosageForm := NULL] %>%
  .[, Form := gsub(";", " (", Form, fixed = T)] %>%
  .[, Form := paste0(Form, ")")]

```

# Save Data

Save result to a `.csv` file.

```{r save-data}

fwrite(mdrugs, file.path(dir, paste0("Off-Patent Drugs - ", nm, ".csv")))

```
